---
title: "Assignment_3"
author: "Ajay Kanubhai Patel"
date: "2022-10-21"
output:
  pdf_document: default
  word_document: default
---

#Loading the package


```{r}
#Load packages to convert file in PDF.

if(!require(tinytex)){install.packages("tinytex")}

```

#This sets the working directory

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'D:/Final Assignment/DATA/Assignment3')

```

This section is for the basic set up.
It will clear all the plots, the console and the workspace.
It also sets the overall format for numbers.

```{r}
if(!is.null(dev.list())) dev.off()
cat("\014") 
rm(list=ls())
options(scipen=9)
```
#To read Excel file in R data frame.


```{r}

if(!require(readxl)){install.packages("readxl")}
library("readxl")
```

#This sets the working directory

```{r} 

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'D:/Final Assignment/DATA/Assignment3')

```


This section is for the basic set up.
It will clear all the plots, the console and the workspace.
It also sets the overall format for numbers.



#To read Excel file in R data frame.


```{r}

if(!require(readxl)){install.packages("readxl")}
library("readxl")
```


```{r}
if(!require(pastecs)){install.packages("pastecs")}
library("pastecs")

if(!require(lattice)){install.packages("lattice")}
library("lattice")

if(!require(vcd)){install.packages("vcd")}
library("vcd")

if(!require(HSAUR)){install.packages("HSAUR")}
library("HSAUR")

if(!require(rmarkdown)){install.packages("rmarkdown")}
library("rmarkdown")


if(!require(ggplot2)){install.packages("ggplot2")}
library("ggplot2")


```


```{r}

getwd()  


Assignment03_AP <- read_excel("PROG8430_Assign03_22F.xlsx")


Assignment03_AP  <- as.data.frame(Assignment03_AP ) 

```  
  
1. 

Initial Transformation

a. Rename all variables with your initials appended.
  
  # to change all column name by appending my initials (Ajay Patel = AP) and 
    separate it by "_".
  # head(data,n) displays 1st n rows present in our excel file. 
    head(Assignment03_BDSA,08)=   shows first 8 rows. If number is not
    provided by default is shows 1st 6 rows

```{r}

colnames(Assignment03_AP) <- paste(colnames(Assignment03_AP), "AP", sep = "_") 


head(Assignment03_AP)

```
b. Transform character variables to factor variables.

  # first we analyze structure of our data and then perform required operation.
  
  
```{r}
str(Assignment03_AP)
```

# to transform chr variables into factor variables.
# There are total 6 varibles in character form.


```{r}

Assignment03_AP$gender_AP <- as.factor(Assignment03_AP$gender_AP)
Assignment03_AP$HR_AP <- as.factor(Assignment03_AP$HR_AP)
Assignment03_AP$BP_AP <- as.factor(Assignment03_AP$BP_AP)
Assignment03_AP$Smoke_AP <- as.factor(Assignment03_AP$Smoke_AP)
Assignment03_AP$Drink_AP <- as.factor(Assignment03_AP$Drink_AP)
Assignment03_AP$Group_AP <- as.factor(Assignment03_AP$Group_AP)
str(Assignment03_AP)

```
Q1 (2). 

Reduce Dimensionality

a. Apply the Missing Value Filter to remove appropriate columns of data.

# To identify missing values we apply summary()func and observe NA's for 
#each column.

#There are 492 records (almost 98.4%) in Income_AP which are NA(Not Available).

#To drop Income_AP column Assignment03_AP[-c(13)].
  

```{r}
summary(Assignment03_AP)

Assignment03_AP <- Assignment03_AP[-c(13)]

head(Assignment03_AP) 


```


b. Apply the Low Variance Filter to remove appropriate columns of data.

#only for numerical columns.

```{r}
stat.desc(Assignment03_AP) #Consider coef of var for the low variance,
summary(Assignment03_AP)

# we can observe that  Hgt_AP has low variance.

table(Assignment03_AP$Hgt_AP) # It displays how many records has same value
# For example, 9 records has 67.1 value in Hgt_AP column.

Assignment03_AP <- Assignment03_AP[-c(8)] #removes Hgt_AP column.

head(Assignment03_AP) 



```


c. Apply the High Correlation Filter to remove appropriate columns of data.

#High correlation between two variables means they have similar trends and are 
#likely to carry similar information.

#No correlation available between numerical and nominal columns.

#pearson, spearman,kendall methos can be used to measure the degree of 
#association between two variables.

# can only check for numerical and we have 4 column with numeric data so
#n(n-1)/2 (4*3/2 = 6) combination should be checked.

# I have checked by three methods just for knowledge.

```{r}

#by spearman method
#Speaman is non-parametric and therefore makes no normalacy assumption

cor(Assignment03_AP$Wgt1_AP,Assignment03_AP$Wgt2_AP,method = "spearman") 
cor(Assignment03_AP$Wgt1_AP,Assignment03_AP$Exercise_AP,method = "spearman") 
cor(Assignment03_AP$Wgt1_AP,Assignment03_AP$WBC_AP,method = "spearman")
cor(Assignment03_AP$Wgt2_AP,Assignment03_AP$Exercise_AP,method = "spearman") 
cor(Assignment03_AP$Wgt2_AP,Assignment03_AP$WBC_AP,method = "spearman") 
cor(Assignment03_AP$Exercise_AP,Assignment03_AP$WBC_AP,method = "spearman")

#by pearson method
#assumes normalacy

cor(Assignment03_AP$Wgt1_AP,Assignment03_AP$Wgt2_AP,method = "pearson") 
cor(Assignment03_AP$Wgt1_AP,Assignment03_AP$Exercise_AP,method = "pearson") 
cor(Assignment03_AP$Wgt1_AP,Assignment03_AP$WBC_AP,method = "pearson")
cor(Assignment03_AP$Wgt2_AP,Assignment03_AP$Exercise_AP,method = "pearson") 
cor(Assignment03_AP$Wgt2_AP,Assignment03_AP$WBC_AP,method = "pearson") 
cor(Assignment03_AP$Exercise_AP,Assignment03_AP$WBC_AP,method = "pearson")

# by kendall method
#Kendall rank correlation (non-parametric) is an alternative to Pearson's 
#correlation (parametric)

cor(Assignment03_AP$Wgt1_AP,Assignment03_AP$Wgt2_AP,method = "kendall") 
cor(Assignment03_AP$Wgt1_AP,Assignment03_AP$Exercise_AP,method = "kendall") 
cor(Assignment03_AP$Wgt1_AP,Assignment03_AP$WBC_AP,method = "kendall")
cor(Assignment03_AP$Wgt2_AP,Assignment03_AP$Exercise_AP,method = "kendall") 
cor(Assignment03_AP$Wgt2_AP,Assignment03_AP$WBC_AP,method = "kendall") 
cor(Assignment03_AP$Exercise_AP,Assignment03_AP$WBC_AP,method = "kendall")


```
#Wgt1_AP and Wgt2_AP are highly correlated to each other so going to drop
# Wgt2_AP column

```{r}

Assignment03_AP <- Assignment03_AP[-c(6)]
head(Assignment03_AP)

```


d. Drop any variables that do not contribute any useful analytical information 
at all.

Answer: Here, I am going to drop gender_AP columns as it contains only Females
so it is not quite useful for analytic purpose.

```{r}

Assignment03_AP <- Assignment03_AP[-c(2)]
head(Assignment03_AP)

```

Q1 (3). 

Outliers

a. Create boxplots of all relevant variables (i.e. numeric, non-binary) 
to determine outliers.

```{r}

str(Assignment03_AP)
boxplot(Assignment03_AP$Wgt1_AP,
       main="Box Plot of Patient Weight 1 week before test start",
        xlab="Weight",
        col="blue", horizontal=TRUE, pch=20)

# let's shrink the graph and observe

boxplot(Assignment03_AP$Wgt1_AP,
       main="Box Plot of Patient Weight 1 week before test start",
        xlab="Weight",
        col="blue", horizontal=TRUE, pch=20, range = 0.5)


boxplot(Assignment03_AP$Exercise_AP, 
        main="Box Plot of Minutes per week patient exercises",
        xlab="Minutes Per Week",
        col="red", horizontal=TRUE, pch=21)

# let's shrink the graph and observe

boxplot(Assignment03_AP$Exercise_AP, 
        main="Box Plot of Minutes per week patient exercises",
        xlab="Minutes Per Week",
        col="red", horizontal=TRUE, pch=21, range = 1)

boxplot(Assignment03_AP$WBC_AP, 
        main="Box Plot of WBC (White Blood Cell)",
        xlab="White Blood Cell Count",
        col="brown",horizontal=TRUE, pch=22)

# let's shrink the graph and observe

boxplot(Assignment03_AP$WBC_AP, 
        main="Box Plot of WBC (White Blood Cell)",
        xlab="White Blood Cell Count",
        col="brown",horizontal=TRUE, pch=22, range = 1)



densityplot( ~ Assignment03_AP$Wgt1_AP, pch=6)
densityplot( ~ Assignment03_AP$Exercise_AP, pch=8)
densityplot( ~ Assignment03_AP$WBC_AP, pch=8)



```

# For factor variables.(for non-binary variables only)
 

```{r}


Assignment03_AP$HR_AP <- as.factor(Assignment03_AP$HR_AP)
Assignment03_AP$BP_AP <- as.factor(Assignment03_AP$BP_AP)


barplot(table(Assignment03_AP$HR_AP), cex.names=0.5, main="Barplot of HR_AP", 
        col = 1)
barplot(table(Assignment03_AP$BP_AP), cex.names=0.5, main="Barplot of BP_AP", 
        col = 2)





```



b. Comment on any outliers you see and deal with them appropriately.

Conclusion:1. wgt1_AP looks good.
           2. Exercise_AP looks good.
           3. WBC_AP has one outlier.
              low HR_AP and low BP_AP categories as comparatively small values
              than that of normal category still they are useful.

#To remove a outlier of WBC_AP at its max value.

```{r}

densityplot( ~ Assignment03_AP$WBC_AP, pch=10)
nr <- which(Assignment03_AP$WBC_AP == max(Assignment03_AP$WBC_AP))  
#above code is to detect Row number with max value 
Assignment03_AP <- Assignment03_AP[-c(nr),]
densityplot( ~ Assignment03_AP$WBC_AP, pch=21)

```

Q2.

Organizing Data

1. Scatter Plots

  a. Create a histogram for one of the Weight variables.
  
```{r}

hist(Assignment03_AP$Wgt1_AP,
     main = "Histogram Of Patient Weight 1 week before test start",
     xlab = "Weight Of Patient",
     xlim = c(45,250),
     ylim = c(0,200),
     col = 4, border = "black")
```
  
  
  
  b. Create a histogram for Exercise.
  
```{r}

hist(Assignment03_AP$Exercise_AP,
     main = "Histogram Of Minutes per week patient exercises",
     xlab = "Minutes Per Week",
     xlim = c(60,350),
     ylim = c(0,120),
     col = 7, border="red")

```
  
  c. Create a scatter plot showing the relationship between Exercise and Weight.
     (note: Exercise should be on the x-axis, Weight should be the y-axis)
     
     
```{r}

plot(Assignment03_AP$Wgt1_AP~Assignment03_AP$Exercise_AP,
     main = "Exercise Vs Weight",
     xlab = "Minutes per week patient exercise",
     ylab =  "Weight Of Patient",
     xlim = c(60,350), 
     ylim = c(45,250),
     col = 2)
```
     
  d. What conclusions, if any, can you draw from the chart?
     From the above chart (scatter plot), we can see that there is not any 
     linear relation between these two variables.
  
--------------------------------------------------------------------------------  
  
  e. Calculate a correlation coefficient between these two variables
     (Wgt1_AP & Exercise_AP)
     What conclusion you draw from it?
     
```{r}

ggplot(data = Assignment03_AP, aes(x = Exercise_AP, 
                                   y = Wgt1_AP, 
                                   color = 7)) +
 geom_point()+ labs(y = "Weight Of Patient",
x = "Minutes per week patient exercise")

xyplot(Wgt1_AP~Exercise_AP, data=Assignment03_AP, main="Exercise  vs Weight", 
       colors=TRUE)

cor(Assignment03_AP$Wgt1_AP,Assignment03_AP$Exercise_AP)
cor.test(Assignment03_AP$Wgt1_AP,Assignment03_AP$Exercise_AP,
         method="spearman")
```


 Conclusion: There is no linear relationship as our 0.00 â‰¤ |ð‘Ÿ| < 0.25.
 
--------------------------------------------------------------------------------
 
Q3 Inference

1. Normality

  a. Create a QQ Normal plot of White Blood Cell counts.

```{r}

qqnorm(Assignment03_AP$WBC_AP, pch=4, frame=TRUE)
qqline(Assignment03_AP$WBC_AP, col=4, lwd=4)

```

  b. Conduct a statistical test for normality on White Blood Cell counts.
  
```{r}

shapiro.test(Assignment03_AP$WBC_AP) #Shapiro-Wilk Normality Test


```
  
c. Is White Blood Cell count normally distributed? What led you to this 
   conclusion?
   
   Yes, White Blood Cell Count is normally distributed. 
   
   1.as p-value = 0.6 which is GREATER THAN 0.05 (P>0.05) that 
     means we cannot reject Null hypothesis which state that variable is 
     normally distributed.
     
   2.From QQ Normal Plot and qq line all the points are on and close to line
     which indicates normality of variable.(from Q3 a)
     
   
-------------------------------------------------------------------------------     
     
Statistically Significant Differences

a. Compare White Blood Cell counts between the treatment and control group 
   using a suitable hypothesis test.
   
```{r}

# to perform F test to see whether variance are equal or not.

Ftest_AP <- var.test(WBC_AP~Group_AP, data = Assignment03_AP)
Ftest_AP

# from F-test p > 0.05 that means no significant difference in variances or
# variance of variables are almost same.

# I have described reasons behind choosing this test below.

Ttest_AP <- t.test(WBC_AP~Group_AP, data = Assignment03_AP, var.equal=TRUE)
Ttest_AP


```
   
   
b. Explain why you chose the test you did.

   a. Data is Independent
   b. Data is normally distributed (From Shapiro-Wilks Test)
   c. Variance is unknown, but equal (From F-Test)
   
   A t-test is a statistical test that is used to compare the means of two 
   groups. It is often used in hypothesis testing to determine whether 
   a process or treatment actually has an effect on the population of 
   interest, or whether two groups are different from one another.
   
   Reference:Bevans, R. (2022, July 9). An introduction to t-tests. 
             Scribbr. Retrieved October 20, 2022, from 
             #https://www.scribbr.com/statistics/t-test/
   
--------------------------------------------------------------------------------   
   
c. Do you have strong evidence that White Blood Cell counts are different 
   between the treatment and control groups?
   
   Yes, From the T-Test, I can say that White Blood Cell Counts are different 
   between the treatment and control groups as p-value is less than 0.05 which 
   provide evidence against Null Hypothesis.
   
   
--------------------------------------------------------------------------------


3. Multiple Statistical Differences

  a. Determine if White Blood Cell count varies by Heart Rate Level using ANOVA
     (statistical) and a sequence of boxplots (graphical).
     
```{r}

summary(aov(WBC_AP~HR_AP, data = Assignment03_AP))

boxplot(WBC_AP~HR_AP, data = Assignment03_AP,
        main = "Heart Rate VS White Blood Cell",
        ylab = "Heart Rate",
        xlab = "White Blood Cell", horizontal = TRUE,
        range = 1)


```

Conclusion: No,  White Blood Cell count do not vary by Heart Rate Level as
            p-value is greater than 0.05 so mean of all the groups is almost
            same.



  b. Determine if White Blood Cell count varies by Blood Pressure Level using 
     ANOVA and a sequence of boxplots.
     
```{r}
summary(aov(WBC_AP~BP_AP, data = Assignment03_AP))

boxplot(WBC_AP~BP_AP, data = Assignment03_AP,
        main = "White Blood Cell vs Blood Pressure",
        ylab = "Blood Pressure",
        xlab = "White Blood Cell", horizontal = TRUE,
        range = 1)
```


Conclusion: No, White Blood Cell count do not vary by Blood Pressure Level as 
             we can see p-value derived from ANOVA Test is greater than 0.05 
             which indicates that means of all BP categories are probably same.
             Moreover, Boxplot also indicates same result.
   
   
   
   
 
 
 





